Plots

### 2021-07-12 plots.Rmd
Added point labels to average misclassification rate vs. cosine similarity to FFPE scatterplot, started working on AUC vs cosine similarity to FFPE scatterplot

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NMF)
library(tidyverse)
library(ROCR)
source("bayes_function.R")
source("get_classification_df.R")
```

```{r}
#load FFPE signature as a matrix

##Function to load FFPE signature
load_old_mutational_matrix<- function(filename,mut.order="supplied_results/mut_sig.order.txt") {

  mutational.signature = read_tsv(filename)
  mut_mat = read_tsv(mut.order)

  mut.join = mutational.signature %>%
    inner_join(mut_mat)

  order.tri = unique(mut.join$ext.context)

  mut.join = mut.join %>%
    mutate(ext.context = factor (ext.context, levels = order.tri)) %>%
    select(ext.context,prob)

  mut.matrix=as.matrix(sapply(mut.join[,2:ncol(mut.join)], as.numeric))  
  rownames(mut.matrix) = mut.join$ext.context

  return (t(mut.matrix)[,1:96])
}

ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
```

@@

```{r}
#Load COSMIC signatures (version 3)
cosmic.signatures <- get_known_signatures()

@@

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NMF)
library(tidyverse)
library(ggrepel)
library(ROCR)
source("bayes_function.R")
source("get_classification_df.R")
source("simulation.R")
```

```{r}
#load FFPE signature as a matrix
ffpe.signature <- as.matrix(load_old_mutational_matrix("supplied_results/ffpe.signature.txt"))
```

@@

```{r}
#Load COSMIC signatures (version 3)
cosmic.signatures <- get_known_signatures(muttype = "snv")

@@

#ROC Curve for Sample of 50 MMR mutations and 50 FFPE mutations

@@

#Scatterplot of misclassification rate vs. cosine similarity of COSMIC signature to FFPE signature

```{r}
#Create vector of the signature names
signature.names <- colnames(cosmic.signatures)
```

```{r}
#Create list of signature matrices with assigned rownames
signature.matrix.list <- list()
for (i in 1:60){
  sig <- cosmic.signatures[,i]
  sig.mat <- as.matrix(sig)
  rownames(sig.mat) <- mutations
  colnames(sig.mat) <- signature.names[i]
  signature.matrix.list[[i]] <- sig.mat
}
```

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.misclassification.rates <- c()
average.auc.values <- c()
for (x in 1:60){
  signature.misclass.vec <- c()
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 180 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],180)
    #Create FFPE sample vector of 20 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 20)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    misclassification.rate <- mean(classification.df$misclassification)
    signature.misclass.vec[y] <- misclassification.rate
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df %>%
      mutate(signature.indicator = ifelse(classify == sig.name, 1, 0))
    pred <- prediction(classification.df.2[3], classification.df.2$signature.indicator)
    auc.perf <- performance(pred, measure = "auc")
    auc.value <- auc.perf@y.values[[1]][1]
    signature.auc.vec[y] <- auc.value
  }
  average.signature.misclass <- mean(signature.misclass.vec)
  average.misclassification.rates[x] <- average.signature.misclass
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

```{r}
#Example of loop for SBS5: all mutations are classified as SBS5 -> problem with ROCR
sample.vec <- create_signature_sample_vector(signature.matrix.list[[5]],180)
    #Create FFPE sample vector of 20 mutations
ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 20)
classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[5],"FFPE"), list(signature.matrix.list[[5]], ffpe.signature))
misclassification.rate <- mean(classification.df$misclassification)
signature.misclass.vec[10] <- misclassification.rate
class.df.colnames <- colnames(classification.df)
sig.name <- class.df.colnames[3]
classification.df.2 <- classification.df %>%
  mutate(signature.indicator = ifelse(classify == sig.name, 1, 0))
pred <- prediction(classification.df.2[3], classification.df.2$signature.indicator)
    auc.perf <- performance(pred, measure = "auc")
    auc.value <- auc.perf@y.values[[1]][1]
    signature.auc.vec[10] <- auc.value
```

```{r}
#create vector of cosine similarities to the FFPE signature
ffpe.cosmic.cos.sims <- c()
for (i in 1:60){
  cos.sim <- cos_sim_matrix(signature.matrix.list[[i]], ffpe.signature)
  ffpe.cosmic.cos.sims[i] <- cos.sim[1,1]
}
```

```{r}
cos.sim.misclass.data <- data.frame(signature.names, ffpe.cosmic.cos.sims, average.misclassification.rates)
colnames(cos.sim.misclass.data) <- c("signature", "cosine_similarity", "average_misclassification")
```

```{r}
ggplot(cos.sim.misclass.data, aes(x = cosine_similarity, y = average_misclassification))+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "Average misclassification rate vs. cosine similarity to FFPE", x = "Cosine similarity of COSMIC signature to FFPE signature", y = "Average misclassification rate (10 iterations)")+
  scale_x_log10()+
  geom_label_repel(aes(label = signature), max.overlaps = 7, label.size =   0.1)
```
